Reproducibility and Open Science in Economics

### 2025-03-06 article.tex
Updates from Overleaf

As a graduate student in economics at Université de Montréal, reading the economics literature was  easy. While the main university library had all the relevant subscriptions, our department librarian, Fethy Mili, would populate the library of the economics department with multi-hued rows of working papers. Mili was also one of the key creators of what was initially known as \ac{WoPEc} and BibEc (for printed working papers) \citep{krichel_wopec_1997,cruz_cataloging_2000,krichel_economics_2009}, populating the latter since 1993 \citep[][p. 450]{batizlazo_brief_2012}. The overall network, known as \ac{RePEc}, was born contemporaneously with the more widely known arXiv and the more centralized \ac{SSRN}. While electronic working papers were mostly free in those days (there was no way to pay for them), Mili's work consisted of sending out postage-paid envelopes to all of the various economics departments that were publishing the working papers, and then cataloging the incoming printed materials electronically, for public consumption. In the end, information about the existence of the working papers was freely available, but access to (printed) working papers still required a small fee to cover the cost of shipping.

@@

Less broadly, \cite{vicente-saez_open_2018} identify a consensus that defines open science as ``transparent and accessible knowledge that is shared and developed through collaborative networks.''
%
In this article, I will focus on the what \cite{unesco_understanding_2022} calls open science ``knowledge'' and will briefly discuss ``infrastructures.'' I will highlight how some elements have been quite widespread in economics for some time. I will try to identify limits to fully open accessibility, some of which are intrinsic to the nature of the research conducted in economics, and describe how widespread such limitations may be. In particular, I will highlight how those access restrictions are not, as many think, an impediment to \textbf{open} science, in the sense that aforementioned ``collaborative networks'' can still access these resources.

A key ambiguity will arise in how big such networks need to be in order to be considered ``open.'' Clearly, $n=2$ is not considered a network. The \ac{NBER} defines its affiliated scholars as a network: $n=1804$ as of January 2025, primarily in North America \citep{national_bureau_of_economic_research_affiliated_2025}, but X authors have published Y NBER working papers in 2024 (SOURCE NEEDED). J-PAL has approximately $n=1725$, at 120 universities on all populated continents \citep{abdul_latif_jameel_poverty_action_lab_affiliated_2025}. Between 2001 and June 2024, there had been $n=2023$ researchers on projects that used confidential U.S. Census Bureau in the \ac{FSRDC} \citep{us_census_bureau_uscensusbureaufsrdc-external-census-projects_2024}. In an average year (2013-2023), $n=\economicsgrads{}$ students graduate from a U.S. university with a Ph.D. in economics \citep[Table 1-5]{national_science_foundation_doctorate_2024}. In the approximately 30 years since inception of \ac{RePEc}, $n=526$ authors from 52 institutions in \textbf{Norway} have published a paper listed on \ac{RePEc} (presumably in economics) \citep{ideasrepec_within_2025}. All of these are measured across different spatio-temporal dimensions. Are they large? Context and purpose matter. Some may intersect. How many U.S. graduate students in the past 10 years have published an NBER working paper and are now at a Norwegian institution?

@@

\paragraph{Organic data:} However, most economists, when asked about data subject to restrictions, will think of ``proprietary data,'' a term often applied to any data that may be subject to restrictions of use and re-use. Access to most administrative data is typically not ``open'' in the sense of OKD, but are they open enough, given the privacy concerns that are attached to these data? Many have argued that access is not broad enough \citep{card_expanding_2010,einav_economics_2014}, while acknowledging the difficulty of addressing privacy and security of the data at scale. \citet{abowd_economic_2019} discuss the challenge of making the choice of between accuracy of (public) statistics and data, and the privacy loss inherent in doing so. \citet{nagaraj_improving_2020} argue that more openness improves scientific progress, and \citet{nagaraj_how_2023} study this in the context of the US system for providing access (\ac{FSRDC}). They note that 4\% of US-based empirical authors have had some access to the FSRDC system. I am not aware of similar studies for other countries, such as France \citep[the equivalent system is the \ac{CASD}, ][]{gadouche_centre_2019} or Canada \citep{currie2015social}. In absolute terms, these networks host several hundred researchers every year. As per (CENSUS), 2,084 researchers have had access to projects involving Census Bureau data since (YEAR).\footnote{\url{https://labordynamicsinstitute.github.io/fsrdc-external-census-projects/}}. \citet{nagaraj_how_2023} mention 861 papers. Similar numbers can be obtained for the French (xxx researchers and 417 publications)
% https://www.casd.eu/toutes-les-publications/liste/0/20/ 
% 20 pages à 20 publications et 1 à 17 = 417
and Canadian (2201  active researchers as of January 2025, and 3245 papers published between 2000 and May 2024)\footnote{Provided by Grant Gibson, CRDCN, on February 10, 2025.} networks.\footnote{The \citet{nagaraj_how_2023} number only includes papers published by economists. All other numbers are counts of researchers and publications in all disciplines, and include non-economists.} The number of publications from access to these networks is smaller in absolute terms then those from PSID and SOEP, though likely higher in impact \citep{nagaraj_how_2023}.

@@

>>>>

As a graduate student in economics at Université de Montréal, reading the economics literature was  easy. While the main university library had all the relevant subscriptions, our department librarian, Fethy Mili, would populate the library of the economics department with multi-hued rows of working papers. Mili was also one of the key creators of what was initially known as \ac{WoPEc} and BibEc (for printed working papers) \citep{krichel_wopec_1997,cruz_cataloging_2000,krichel_economics_2009}, populating the latter since 1993 \citep[][p. 450]{batizlazo_brief_2012}. The overall network, known as \ac{RePEc}, was born contemporaneously with the more widely known arXiv \citep{ginsparg_it_2011} and the more centralized \ac{SSRN} \citep{noauthor_social_2025}. While electronic working papers were mostly free in those days (there was no way to pay for them), Mili's work consisted of sending out postage-paid envelopes to all of the various economics departments that were publishing the working papers, and then cataloging the incoming printed materials electronically, for public consumption. In the end, information about the existence of the working papers was freely available, but access to (printed) working papers still required a small fee to cover the cost of shipping.

@@

Less broadly, \citet{vicente-saez_open_2018} identify a consensus that defines open science as ``transparent and accessible knowledge that is shared and developed through collaborative networks.''
%
In this article, I will focus on the what \cite{unesco_understanding_2022} calls open science ``knowledge'' and will briefly discuss ``infrastructures.'' I will highlight how some elements have been quite widespread in economics for some time. I will try to identify limits to fully open accessibility, some of which are intrinsic to the nature of the research conducted in economics, and describe how widespread such limitations may be. In particular, I will highlight how those access restrictions are not, as many think, an impediment to \textbf{open} science, in the sense that aforementioned ``collaborative networks'' can still access these resources.

A key ambiguity will arise in how big such networks need to be in order to be considered ``open.'' Clearly, $n=2$ is not considered a network. The \ac{NBER} defines its affiliated scholars as a network: $n=1804$ as of January 2025, primarily in North America \citep{national_bureau_of_economic_research_affiliated_2025}. However, in \nberyear{}, a total of \nberauthorsunique{} authors published \nberpapers{} NBER working papers. J-PAL has approximately $n=1725$, at 120 universities on all populated continents \citep{abdul_latif_jameel_poverty_action_lab_affiliated_2025}. Between 2001 and June 2024, there had been $n=2023$ researchers on projects that used confidential U.S. Census Bureau in the \ac{FSRDC} \citep{us_census_bureau_uscensusbureaufsrdc-external-census-projects_2024}. In an average year (2013-2023), $n=\economicsgrads{}$ students graduate from a U.S. university with a Ph.D. in economics \citep[Table 1-5]{national_science_foundation_doctorate_2024}. In the approximately 30 years since inception of \ac{RePEc}, $n=526$ authors from 52 institutions in \textbf{Norway} have published a paper listed on \ac{RePEc} (presumably in economics) \citep{ideasrepec_within_2025}. All of these are measured across different spatio-temporal dimensions. Are they large? Context and purpose matter. Some may intersect. How many U.S. graduate students in the past 10 years have published an NBER working paper and are now at a Norwegian institution?

@@

\paragraph{Organic data:} However, most economists, when asked about data subject to restrictions, will think of ``proprietary data,'' a term often applied to any data that may be subject to restrictions of use and re-use. Access to most administrative data is typically not ``open'' in the sense of OKD, but are they open enough, given the privacy concerns that are attached to these data? Many have argued that access is not broad enough \citep{card_expanding_2010,einav_economics_2014}, while acknowledging the difficulty of addressing privacy and security of the data at scale. \citet{abowd_economic_2018} discuss the challenge of making the choice of between accuracy of (public) statistics and data, and the privacy loss inherent in doing so. \citet{nagaraj_improving_2020} argue that more openness improves scientific progress, and \citet{nagaraj_how_2023} study this in the context of the US system for providing access (\ac{FSRDC}). They note that 4\% of US-based empirical authors have had some access to the FSRDC system. I am not aware of similar studies for other countries, such as France \citep[the equivalent system is the \ac{CASD}, ][]{gadouche_centre_2019} or Canada \citep{currie2015social}. In absolute terms, these networks host several hundred researchers every year. As per (CENSUS), 2,084 researchers have had access to projects involving Census Bureau data since (YEAR).\footnote{\url{https://labordynamicsinstitute.github.io/fsrdc-external-census-projects/}}. \citet{nagaraj_how_2023} mention 861 papers. Similar numbers can be obtained for the French (xxx researchers and 417 publications)
% https://www.casd.eu/toutes-les-publications/liste/0/20/ 
% 20 pages à 20 publications et 1 à 17 = 417
and Canadian (2201  active researchers as of January 2025, and 3245 papers published between 2000 and May 2024)\footnote{Provided by Grant Gibson, CRDCN, on February 10, 2025.} networks.\footnote{The \citet{nagaraj_how_2023} number only includes papers published by economists. All other numbers are counts of researchers and publications in all disciplines, and include non-economists.} The number of publications from access to these networks is smaller in absolute terms then those from PSID and SOEP, though likely higher in impact \citep{nagaraj_how_2023}.

@@

% also Goldsmith-Pinkham, Paul, Hull, Peter, and Kolesar, Michal. Data and Code for: Contamination Bias in Linear Regressions. Nashville, TN: American Economic Association [publisher], 2025. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2025-01-03. https://doi.org/10.3886/E207983V1
%
\begin{quote}
    ``These studies were identified by a systematic search of papers in the AEA Data and Code Repository''
\end{quote}
